%-----------------------------------------------------------------------------
%  Bernoulli-EKF Multi-Target Tracker on SE(3) — design + implementation
%  Four-part document: (I) compact algorithmic summary, (II) detailed
%  algorithmic exposition (no code references), (III) algorithm-to-code
%  map (one file path per line), (IV) reviewer-style critique.
%-----------------------------------------------------------------------------
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm,enumitem,hyperref}
\usepackage{algpseudocode,algorithm}
\usepackage{xcolor}

% Verdict tags used in Part V.
\newcommand{\REDUNDANT}{\textcolor{red}{\textbf{REDUNDANT.}}}
\newcommand{\REFERENCE}{\textbf{Reference.}}
\newcommand{\PRINCIPLED}{\textbf{Principled.}}

\DeclareMathOperator{\Tr}{tr}
\DeclareMathOperator{\Exp}{Exp}
\DeclareMathOperator{\Log}{Log}
\newcommand{\R}{\mathbb{R}}
\newcommand{\SE}{\mathrm{SE}(3)}
\newcommand{\so}{\mathrm{SO}(3)}
\newcommand{\se}{\mathfrak{se}(3)}
\newcommand{\Ad}{\mathrm{Ad}}
\newcommand{\file}[1]{\texttt{#1}}

\title{Bernoulli-EKF on $\SE$:\\
  A Compact Multi-Object Tracker for Mobile Manipulation}
\author{}
\date{}

\begin{document}
\maketitle

%=============================================================================
\section*{Part I --- One-page Algorithmic Summary}
\label{sec:summary}

The system maintains per-object Bernoulli--Gaussian beliefs over rigid-body
poses on $\SE$, conditioned on RGB-D detections from a moving camera mounted
on a manipulator. Each frame $k$ ingests the SLAM base pose $T_{wb,k}$, the
camera-to-base extrinsic $T_{bc,k}$, the gripper-to-base $T_{bg,k}$, the
gripper finger width, and a list of detections
$\{(\mathrm{pid}_d,\mathrm{label}_d,\mathrm{mask}_d,\mathrm{score}_d)\}_d$
projected through depth.

\paragraph{1. State.} Track $o$ stores
$(\mu_{bo},\,\Sigma_{bo},\,r_o,\,\ell_o,\,\tau_o)$:
$\mu_{bo}\in\SE$ is the object pose in the \emph{base} frame, $\Sigma_{bo}\in
\R^{6\times 6}$ is the body-frame tangent covariance at $\mu_{bo}$, $r_o\in
[0,1]$ is the existence probability, $\ell_o$ a label histogram, and
$\tau_o$ the SAM2 perception id. \emph{Base-frame storage is the canonical
choice}: SLAM uncertainty $\Sigma_{wb}$ never enters the recursion; the world
view is the deterministic composition $T_{wo}=T_{wb}\,\mu_{bo}$.

\paragraph{2. Predict.} Two motion regimes apply per track. \emph{World-static}:
$\mu_{bo,k}=u_k\,\mu_{bo,k-1}$ with $u_k=T_{wb,k}^{-1}T_{wb,k-1}$, propagated
through the adjoint with phase-scheduled $Q_{\text{static}}$. For tracks
unobserved at $k{-}1$ the covariance update is \emph{skipped}, freezing
$\Sigma_{bo}$ at the last-observed keyframe; combined with the deterministic
mean update this preserves $T_{wo,k}=T_{wb,k_f}\,\mu_{bo,k_f}$ for static
objects under any base motion. \emph{Rigid-attached}: held tracks evolve as
$\mu_{bo,k}=\delta_{bg,k}\,\mu_{bo,k-1}$ with phase-aware $Q_{\text{manip}}$.
At the holding$\to$idle (release) transition, a one-shot gravity-aware predict
overwrites the just-released track's world-frame pose and covariance using
voxel-occupancy-driven drop height and a per-label restitution / footprint
table.

\paragraph{3. Detection conditioning.} Subpart suppression voxelises masks
through depth at 2~cm and absorbs detections whose voxel sets are mostly
contained in another's; centroid back-projection drops masks without valid
depth.

\paragraph{4. Association.} Translation-only Mahalanobis with a $\chi^2_3$
gate at $G_{\text{out},t}{=}21.108$ and $R_{cam}{=}(0.02\,\mathrm{m})^2 I$,
plus a hard Euclidean residual cap (default 0.30~m) that catches inflated
covariances. Hungarian on the gated cost; a SAM2-id continuity bonus
$\alpha{=}4.4$ (in $\chi^2$ units) breaks ties. A configurable label gate
runs in either hard mode (mismatched labels become infeasible) or soft mode
(label mismatch penalty $\beta$ + score weight $\gamma$).

\paragraph{5. Update.} For matched pairs on the held track, two safety
prefilters guard the Joseph step: a held-prefilter ($\le\!0.25$~m from the
end-effector) and an innovation clamp ($\le\!0.20$~m from $\mu_w$). The
Joseph step runs on the \emph{translation block only}; the rotation block and
cross-covariance are zeroed (a translation-only observation carries no
rotation information). The world-frame view used by all downstream consumers
is the canonical composition $T_{wo}=T_{wb}\,\mu_{bo}$ with
$\Sigma_{wo}=\Ad(\mu_{bo}^{-1})\,\Sigma_{wb}\,\Ad(\mu_{bo}^{-1})^\top
+\Sigma_{bo}$.

\paragraph{6. Birth and prune.} A confirm-3 / score-floor / border-margin /
near-live admission chain controls births. Survivors mint a fresh oid, run
ICP once at admission, and seed $r_0=\lambda_b\hat s_d/(\lambda_b\hat s_d
+\lambda_c)$. Tracks prune when $r_o<10^{-3}$, except the held seed is
floored at $r{=}0.5$.

\paragraph{7. Self-merge.} Same-label tracks within 5~cm fuse via information
sum, except pairs in the relation-protected set (the scene graph asserts
they are physically distinct). The held seed is always the keeper.

\paragraph{8. Relations and held-set expansion.} A VLM/REST backend computes
in/on edges at first call, release transitions, new-oid arrivals, and every
90 frames. The parent matrix is converted to edges with the convention
\textbf{parent~=~supported, child~=~supporter}; a per-edge EMA
($\alpha{=}0.3$, threshold $0.5$) smooths flicker. The held set is the
closure of $\{\text{held seed}\}$ under in/on edges; after each self-merge,
EMA keys are remapped (drop $\to$ keep) so the closure stays valid.

\bigskip\hrule\bigskip

%=============================================================================
\section*{Part II --- Detailed Algorithmic Exposition}
\label{sec:algo}

This part is purely algorithmic; it carries no source-file references. The
algorithm-to-code map is in Part~III.

\subsection*{II.1 State}

$$\mathcal{S}_o = (\mu_{bo}\in\SE,\ \Sigma_{bo}\in\R^{6\times 6}_{\succeq 0},\
  r_o\in[0,1],\ \ell_o,\ \tau_o\in\mathbb{Z},\ \text{age},\ \text{frames since obs}).$$

Body-frame tangent at $\mu_{bo}$: a perturbation $\xi\in\se$ acts as
$\mu' = \mu\,\Exp(\xi)$ (right-multiply). Ordering
$\xi=(\xi_v;\xi_\omega)$ --- \emph{translation first, then rotation}. The
adjoint
$$\Ad(T) = \begin{bmatrix} R & [t]_\times R \\ 0 & R \end{bmatrix}$$
moves a tangent vector through $T=(R,t)$.

\subsection*{II.2 Predict}

\textbf{World-static.} A world-static object satisfies
$T_{wb,k}\,\mu_{bo,k} = T_{wb,k-1}\,\mu_{bo,k-1}$, hence
$\mu_{bo,k} = u_k\,\mu_{bo,k-1}$ with
$u_k = T_{wb,k}^{-1}T_{wb,k-1}$.

The covariance recursion branches on whether the track was observed at
$k{-}1$:
\begin{equation}\label{eq:proc}
\Sigma_{bo,k} \;=\;
\begin{cases}
\Phi\!\bigl(\Ad(u_k)\,\Sigma_{bo,k-1}\,\Ad(u_k)^\top + Q_{\text{static}}\bigr) & \text{observed at } k{-}1, \\[2pt]
\Sigma_{bo,k-1} & \text{unobserved at } k{-}1 \text{ (frozen keyframe)}.
\end{cases}
\end{equation}
The mean update $\mu_{bo,k}=u_k\mu_{bo,k-1}$ runs in \emph{both} branches,
deterministically. $Q_{\text{static}}$ is scheduled on frames since the
last observation: $1{\rm e}{-}6\,I_6$ (idle, $<5$),
$1{\rm e}{-}5\,I_6$ ($<50$), $1{\rm e}{-}8\,I_6$ thereafter.
$\Phi$ is the optional covariance saturation
\begin{equation}\label{eq:phi}
\Phi(P) \;=\; \min\!\bigl(1,\;\Tr P_{\max}/\Tr P\bigr)\,P,
\end{equation}
inactive when no cap is configured (the default). The Bernoulli predict on
existence is
\begin{equation}\label{eq:bern_pred_r}
r_{k|k-1} = p_s\,r_{k-1|k-1}, \qquad p_s = 1 \text{ default}.
\end{equation}

\paragraph{Why the unobserved branch freezes $\Sigma_{bo}$ (the keyframe
mechanism).}
The Ad-conjugate transport carries a cross-coupling
$[t(u_k)]_\times P_{rr}[t(u_k)]_\times^\top$ that pumps rotation
uncertainty into the translation block each frame. For tracks with high
$\sigma_{yaw}$ (e.g.\ a freshly-released object with $\sigma_{yaw}=\pi$
from the gravity prior, \eqref{eq:gravity_sigma}), the per-frame
contribution $\Delta x^2\!\cdot\!\pi^2$ compounds catastrophically (order
of one metre over a few dozen frames). The keyframe mechanism exploits the
algebraic identity that, telescoping the deterministic mean update over
$K$ frames since the last observation at $k_f$,
\begin{equation}\label{eq:telescope}
\mu_{bo,K} \;=\; u_K \cdots u_{k_f+1}\,\mu_{bo,k_f}
            \;=\; T_{wb,K}^{-1}\,T_{wb,k_f}\,\mu_{bo,k_f}.
\end{equation}
Composing with the current SLAM pose then gives
$T_{wo,K}=T_{wb,K}\mu_{bo,K}=T_{wb,k_f}\mu_{bo,k_f}$ --- \emph{constant} for
a static object, exactly as required. The keyframe is not stored
explicitly; $\Sigma_{bo,K}=\Sigma_{bo,k_f}$ is preserved by skipping the
covariance update. World composition (\S II.5, \eqref{eq:world_compose})
then uses the \emph{current} $\Sigma_{wb,K}$ for the lever-arm
contribution, so the world-frame uncertainty grows only via the SLAM
marginal --- never via per-frame predict drift.

\textbf{Rigid-attached.} For tracks in the held set $\mathcal{H}_k$,
$\mu_{bo,k} = \delta_{bg,k}\,\mu_{bo,k-1}$ with
$\delta_{bg,k} = T_{bg,k}T_{bg,k-1}^{-1}$. A phase-aware $Q_{\text{manip}}$
applies:
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Seed during grasping/releasing: $\mathrm{diag}((0.02)^2{\times}3,\,(0.10)^2{\times}3)$ --- physical grip slack.
\item Seed during holding: $\mathrm{diag}(2.5{\rm e}{-}4{\times}3,\,1{\rm e}{-}3{\times}3)$ --- settled grip.
\item Transitive members of the held set (e.g.\ items resting on a held tray): always the holding row --- they don't deform when the gripper closes around the seed.
\end{itemize}
The held branch runs the full Ad-conjugate (no keyframe-mechanism shortcut),
so the same translation/rotation cross-pumping applies --- bounded in
practice by $\|t(\delta_{bg,k})\|\!\ll\!1$ but \emph{not} eliminated. Edge
cases with $\sigma_{yaw}\!\to\!\pi$ inside the holding window (re-grasps,
release transitions) remain a known caveat.

\subsection*{II.3 Detection conditioning}

\textbf{Subpart suppression.} Voxelise each mask through depth at 2~cm; for
ordered pair $(a,b)$ with $|V_b|\le|V_a|$, if $|V_b\cap V_a|/|V_b|>0.8$,
absorb $b$ into $a$ (label histograms merge). This catches SAM mask splits
and re-prompts.

\textbf{Centroid validity.}
$c_d^{\text{cam}} = \frac{1}{|m_d\cap\text{valid-depth}|}\sum_{p\in m_d}\mathrm{back\_project}(p,\mathrm{depth},K)$.
If the mask has no valid-depth pixels, $d$ is dropped.

\subsection*{II.4 Association}

For each $(o,d)$:
\begin{align*}
  t_{bo}^{\text{meas}} &= T_{bc}\bigl[c_d^{\text{cam}}; 1\bigr]_{1:3} \\
  R_{bo}^t &= R_{bc}\,R_{cam}\,R_{bc}^\top, \quad R_{cam} = \mathrm{diag}((0.02\,\mathrm{m})^2{\times}3) \\
  \nu_{od} &= t_{bo}^{\text{meas}} - \mu_{bo}[1{:}3, 4],\quad
  S_{od} = P_{tt,o} + R_{bo}^t \\
  d^2_{t,od} &= \nu_{od}^\top S_{od}^{-1} \nu_{od}.
\end{align*}

\paragraph{Cost.}
Two modes, selected by a label-enforcement flag:
\begin{equation}\label{eq:cost_sam2}
C_{od} \;=\; d^2_{t,od}
  \;-\;\alpha_{\text{SAM2}}\,\bm 1[\tau_o = \mathrm{pid}_d,\ \tau_o\ge 0]
  \;+\;\beta\,\bm 1[\hat\ell_d \notin \mathcal H_o]
  \;+\;\gamma\,\max(0,\,1-\hat s_d).
\end{equation}
\textbf{Hard label gate} (paper baseline): $\beta=\gamma=0$, and any cell
with mismatched label is forced to a sentinel infeasible value.
\textbf{Soft mode}: $\beta$ is a configurable label penalty (default 0.0;
the production driver sets 6.0) and $\gamma$ is a configurable score weight
(default 0.0; production sets 2.0). $\alpha_{\text{SAM2}}=4.4$ in either
mode.

\paragraph{Label-purity gate (soft mode only).}
The membership test $\hat\ell_d \in \mathcal H_o$ is not pure dictionary
membership; a label is ``meaningfully in the history'' iff it has $\ge\!5$
observations \emph{and} $\ge\!5\%$ of the total. This rejects stale
trace-amount labels (e.g.\ 3 obs against 53\,990) that would otherwise bypass
the $\beta$ penalty.

\paragraph{Outer gates.}
The Hungarian solver minimises $\sum C_{o\sigma(o)}$. Two gates apply:
\begin{enumerate}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Probabilistic $\chi^2_3$: $d^2_{t,od} > G_{\text{out},t}=21.108$
  (i.e.\ at $99.97\%$) $\Rightarrow$ cell infeasible.
\item Hard Euclidean residual cap: $\|\nu_{od}\|_2$ above a configurable
  threshold (default $0.30$~m; \texttt{None} disables) $\Rightarrow$ cell
  infeasible. This catches the inflated-covariance pathology where, after
  a long miss run with $\sigma_t\!\approx\!0.4$~m, an $0.86$~m residual gives
  $d^2_t\!\approx\!17 < 21.108$ and the $\chi^2$ gate becomes meaningless.
\end{enumerate}
\begin{equation}\label{eq:hung}
\sigma^* = \arg\min_\sigma \sum_{o} C_{o\,\sigma(o)},\qquad
\text{drop } (o,d) \text{ if } C_{od} \text{ is infeasible}.
\end{equation}

\subsection*{II.5 Update}

\textbf{Held-track prefilters.} For each accepted match $(o,d)$ where $o$ is
the held seed:
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Held-prefilter: $\|c_d^w - (T_{wb}T_{bg})_{1:3,4}\|_2 \le 0.25$~m.
\item Innovation clamp: $\|c_d^w - (T_{wb}\mu_{bo})_{1:3,4}\|_2 \le 0.20$~m.
\end{itemize}
A failure marks the detection consumed (so it does not re-birth) and skips
the EKF step; the rigid-attach predict carries the track.

\textbf{Joseph step (translation-only).} Let $K_{tt} = P_{tt}\,S^{-1}$
($3\times 3$):
\begin{align*}
  \delta_t^{\text{base}} &= K_{tt}\,\nu \\
  \mu_{bo}[1{:}3, 4] &\leftarrow \mu_{bo}[1{:}3, 4] + \delta_t^{\text{base}} \\
  P_{tt}^{\text{post}} &= (I - K_{tt})\,P_{tt}\,(I - K_{tt})^\top + K_{tt}\,R_{bo}^t\,K_{tt}^\top \\
  \Sigma_{bo}^{\text{post}} &= \mathrm{block\!-\!diag}\bigl(P_{tt}^{\text{post}},\ P_{rr}\bigr).
\end{align*}
The cross-covariance is zeroed: a translation-only observation provides no
information about rotation (we set $R_{rot}=\infty$), so the honest
posterior is block-diagonal. An earlier version preserved the prior
cross-covariance, which broke the Schur condition
$P_{rr}-P_{rt}^\top P_{tt}^{-1}P_{rt}\succeq 0$ over many updates; the
current version is PSD by construction, with an eigenvalue clip as a
belt-and-suspenders.

For pairs where ICP succeeded, an alternative full-$\SE$ update via the
Joseph form on the $6$-D innovation lifts the camera-frame measurement to
base frame:
\begin{equation}\label{eq:meas_compose}
T_{bo}^{\text{meas}} = T_{bc}\,\hat T^{co}, \qquad
R^{bo} = \Ad(T_{bc})\,\hat R^{co}\,\Ad(T_{bc})^\top,
\end{equation}
followed by IEKF inner iterations and Huber inner / outer reweighting
(\eqref{eq:huber_w}). The EKF likelihood used by \eqref{eq:r_assoc} is
\begin{equation}\label{eq:ekf_lik}
\log\mathcal L = -\tfrac{1}{2}\bigl(\nu^\top S^{-1}\nu + \log\det(2\pi S)\bigr),
\end{equation}
computed in log space to avoid underflow at $\log\mathcal L\sim -50$.

\paragraph{World-frame composition.}\label{sec:world_compose}
The canonical world-frame view consumed by every downstream client
(relations, slow tier, gravity-predict input, visualizer) is
\begin{equation}\label{eq:world_compose}
\mu_{wo} = T_{wb}\,\mu_{bo}, \qquad
\Sigma_{wo} = \Ad(\mu_{bo}^{-1})\,\Sigma_{wb}\,\Ad(\mu_{bo}^{-1})^\top
              + \Sigma_{bo}.
\end{equation}
$\Sigma_{wb}$ is read from the current SLAM marginal at query time, so the
world-cov tracks any SLAM drift / loop-closure shift even when no new
object observation has arrived. Combined with the keyframe-mechanism freeze
of $\Sigma_{bo}$ (\eqref{eq:proc}), an unobserved static track's world cov
grows \emph{only} via the $\Sigma_{wb}$ lever --- the user-intended
behaviour.

\paragraph{World-frame overwrite (release hook).}
A pose-overwrite primitive forces the EKF state for one oid from a
world-frame $(T_{wo},P_{wo})$ input. It is invoked at the release transition
to install the gravity-aware landing prior (\eqref{eq:gravity_sigma}); the
body-frame storage becomes $\mu_{bo}=T_{wb}^{-1}T_{wo}$ and
$\Sigma_{bo}=P_{wo}$ (the right-trivialised tangent transport reduces to the
identity at the new $\mu_{bo}$). After the overwrite, the keyframe-mechanism
freeze in \eqref{eq:proc} keeps the gravity-supplied covariance
($\sigma_{yaw}=\pi$, $\sigma_{xy}\!\sim\!5$--$10$~cm) intact across
subsequent unobserved frames.

\paragraph{Per-track observation chain (diagnostic).}\label{sec:chain}
Each track also accumulates an append-only chain
$\mathcal{H}^{(i)} = [(\,k,\hat T^{co}_k,\hat R^{co}_k\,),\ldots]$ of
camera-frame ICP measurements. \textbf{The chain is NOT the canonical
world-frame source} --- that role belongs to the world composition
\eqref{eq:world_compose}. The chain is computed in parallel by the
visualizer for cross-checking and is the natural substrate for a future
loop-closure-aware smoother.

\paragraph{Huber reweight.}
\begin{equation}\label{eq:huber_w}
w(d^2; G_{\text{in}}, G_{\text{out}}) = \begin{cases}
1 & d^2 \le G_{\text{in}}, \\
G_{\text{in}}/d^2 & G_{\text{in}} < d^2 \le G_{\text{out}}, \\
0 & d^2 > G_{\text{out}}.
\end{cases}
\end{equation}
Defaults: $G_{\text{in},t}=7.815$ ($\chi^2_3$ at 95\%),
$G_{\text{out},t}=21.108$.

\subsection*{II.6 Birth and prune}

Birth gates:
\begin{enumerate}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Border: bbox not within $2$~px of the image edge.
\item Confirm-$k$: tracker-side pending counter $\ge k=3$ (counts the
  tracker's unmatched-streak frames, not perception's $n_{\text{obs}}$ which
  resets on SAM2 id reissue).
\item Score floor: rolling-max score $\ge 0.20$.
\item Near-live: world-frame centroid $> 0.05$~m from any same-label live
  track. For the held oid the radius widens to $0.25$~m and $T_{we}$ is
  checked as a secondary anchor.
\item TTL: pending entries that go $30$ frames without a matching arrival
  are evicted (one second at $30$~Hz).
\end{enumerate}
Survivors mint a fresh oid, run ICP once on admission, and seed
\begin{equation}\label{eq:birth_r}
r_0 \;=\; \frac{\lambda_b\,\hat s_d}{\lambda_b\,\hat s_d + \lambda_c}.
\end{equation}
Birth covariance defaults to
\begin{equation}\label{eq:birth_state}
P_{\text{new}} = \hat R^{co}\quad (\text{from ICP, by default})
\end{equation}
and falls back to
$\mathrm{diag}((5\,\text{cm})^2{\times}3,(0.05\,\text{rad})^2{\times}3)$
when the ICP-derived form is disabled.

\paragraph{Existence updates.}
For matched pairs
\begin{equation}\label{eq:r_assoc}
r' = \frac{(1 - p_s) + p_s\,p_d\,\mathcal L\,r}{1 - p_s + p_s\,r},
\end{equation}
and for missed tracks (with $\tilde p_d = p_d\,p_v$)
\begin{equation}\label{eq:r_miss}
r' = \frac{(1 - \tilde p_d)\,r}{1 - \tilde p_d\,r},
\end{equation}
with $p_v$ the visibility probability of \eqref{eq:pv}.

\paragraph{Prune.}
$r_o < 10^{-3}$ deletes the track, EXCEPT the held seed, whose existence is
floored at $0.5$ (survives long held-occlusion).

\subsection*{II.7 Self-merge}

Inputs: the held seed $s$, and protected pairs $\mathcal P$ derived from the
relation edge set $\mathcal E$ (\S II.9). The parent/child convention is
\emph{parent~=~supported, child~=~supporter}, so a relation $(p,c,\text{on})$
contributes the unordered pair $(\min(p,c),\max(p,c))$ to $\mathcal P$:
\begin{equation*}
\mathcal{P} = \bigl\{\,(\min(p,c),\max(p,c))\;:\;(p,c,\cdot)\in\mathcal{E}\,\bigr\}.
\end{equation*}
Same-label pairs in $\mathcal P$ are protected from collapse --- the scene
graph asserts they are physically distinct (item-on-tray, etc.).
Information-sum fusion (with the keep mean as linearisation point) is
\begin{equation}\label{eq:self_merge_info}
\begin{aligned}
P^{-1}_{\text{new}} &= P^{-1}_{\text{keep}} + P^{-1}_{\text{drop}}, \\
\mu_{\text{new}} &= \mu_{\text{keep}}\,\Exp\!\bigl(P_{\text{new}}\,P^{-1}_{\text{drop}}\,\Log(\mu_{\text{keep}}^{-1}\mu_{\text{drop}})\bigr).
\end{aligned}
\end{equation}

\begin{algorithm}[H]
\caption{Self-merge pass.}
\begin{algorithmic}[1]
\For{each pair $(o_i, o_j)$ with $\ell_i = \ell_j$}
  \If{$(\min(o_i,o_j),\max(o_i,o_j))\in\mathcal{P}$} \textbf{continue} \EndIf
  \State $d \gets \|\mu_w(o_i) - \mu_w(o_j)\|_2$
  \If{$d > 0.05$} \textbf{continue} \EndIf
  \State Push $(d, d^2_t, o_i, o_j)$ as candidate.
\EndFor
\State Sort candidates by $d$ ascending.
\For{each $(d, d^2_t, o_i, o_j)$}
  \If{$o_i$ or $o_j$ already absorbed} \textbf{continue} \EndIf
  \State Choose keep$\gets s$ if $s\in\{o_i,o_j\}$, else higher-$r$, else lower oid.
  \State Information-sum fuse $\mu, \Sigma$; merge label histograms / SAM2 id / $T_{oe}$ / $r$.
  \State Append \{keep, drop, $d$, $d^2_t$\} to merges.
\EndFor
\end{algorithmic}
\end{algorithm}
The protected-pairs gate is what stops a same-label item-on-tray pair from
collapsing once their world centroids drift within 5~cm under occlusion
bias.

\subsection*{II.8 Visibility}\label{sec:vis}

For each track $i$, sample object-surface points
$\{x_j^{(i)}\}_{j=1}^M$ on a Fibonacci sphere of radius $r_{\text{obj}}$
around $\mu_{bo}$ (the paper specifies projecting the per-object TSDF
shape summary; the current implementation uses the spherical fallback).
Lift each through $T_{bc}$ into camera, depth-test against the current depth
image:
\begin{equation}\label{eq:pv}
p_v^{(i)} \;=\;
\frac{\#\bigl\{j : x_j^{(i)} \text{ in-frustum and passes depth test}\bigr\}}
     {\#\bigl\{j : x_j^{(i)} \text{ in-frustum}\bigr\}}
\quad\in\;[0,1].
\end{equation}
$p_v=1$ when no in-frustum sample is occluded; $p_v=0$ when fully
out-of-frustum or fully occluded. The continuous probability composes
naturally with \eqref{eq:r_miss}; the four-state machine
(present~/~occluded~/~absent~/~not-in-view) collapses into a single weight
on the miss-frame likelihood.

\subsection*{II.9 Relations}

\textbf{Trigger gate.} The relation backend fires at first call, at release
transitions, on new-oid arrivals, or every 90 frames. The grasp-transition
trigger was removed (the backend at grasp onset has incomplete information).

\textbf{Backend.} On a triggered frame, the orchestrator builds normalised
bboxes per oid (from the detection-to-oid map) and calls a VLM (LLM-based)
or REST relation client. The backend returns $p\in[0,1]^{N\times N}$ with
$p_{ij}$ = ``$j$ rests on $i$''. A disk cache keyed by
$(\text{frame}, N, \text{rounded bboxes})$ is read on retry.

\textbf{Convention swap.} Held-set expansion uses parent~=~supported,
child~=~supporter. The backend's $(i,j)$ encodes the opposite ($j$ rests on
$i$). Edge emission swaps:
$$\mathrm{edge} = \mathrm{RelationEdge}(\text{parent}=j,\text{child}=i,\text{type}=\text{on},\text{score}=p_{ij}).$$

\textbf{EMA.} For each key $k=(\text{parent},\text{child},\text{type})$:
$$\mathrm{ema}_k(t) = \alpha\,\mathrm{raw}_k(t) + (1-\alpha)\,\mathrm{ema}_k(t-1),\quad \alpha=0.3.$$
Edges emit when $\mathrm{ema}\ge 0.5$ and prune when $\mathrm{ema}\le
0.01$. Half-life is $\sim\!2$ trigger calls $\sim\!180$ frames at 30~Hz.

\textbf{Remap after merges.} When self-merge produces
$\{(\text{drop}_i,\text{keep}_i)\}$, every EMA key and emitted edge has
parent/child rewritten via the closure of \texttt{drop\_to\_keep}.
Self-loops drop, duplicate keys merge by max.

\subsection*{II.10 Held-set expansion}\label{sec:held}

\begin{algorithm}[H]
\begin{algorithmic}[1]
\State $S \gets \{s\}$
\For{$\textrm{iter}=1\ldots 8$}
  \State changed $\gets$ false
  \For{each edge $(p,c,t)\in\mathcal{E},\ t\in\{\text{in},\text{on}\}$}
    \If{$c\in S$ and $p\notin S$} $S \gets S\cup\{p\}$; changed$\gets$true \EndIf
  \EndFor
  \If{not changed} \textbf{break} \EndIf
\EndFor
\State \textbf{return} $S$
\end{algorithmic}
\end{algorithm}
The caller post-filters $S$ against currently-live oids to drop references
to pruned tracks the EMA has not yet decayed.

\subsection*{II.11 Gravity-aware predict at release}\label{sec:gravity}

When the gripper releases an object, the rigid-attachment predict no longer
applies and the static predict over-states the prior (the object physically
falls). This subsystem replaces the EKF mean and covariance for the
just-released oid with the post-settling prediction in a single step at the
$\textsc{Releasing}\!\to\!\textsc{Idle}$ transition.

\paragraph{Voxel observability.}
A dense 3-state grid (\textsc{Unseen}\,/\,\textsc{Empty}\,/\,\textsc{Occupied})
plus per-voxel hit/pass counters with hysteresis ($\ge\!2$ confirming hits
and $\ge\!3$ pass-through rays before transition) is updated each frame
from depth: each pixel's ray marks voxels traversed before the depth hit
as pass-pending and the voxel at the hit as hit-pending. Default workspace
$5\,\text{m}\times 5\,\text{m}\times 3\,\text{m}$ at $5\,\text{cm}$ voxel
size.

\paragraph{Object dynamics property.}
A per-label table maps each detection label to
$(e,\,\mu,\,\text{shape},\,r_{\text{obj}})$. Restitution defaults from
agricultural / engineering literature; a generic fallback with
$e=0.40,\,r=0.05$ applies to unknown labels. A perception-side override
hook lets a learned estimator supersede the table per-detection.

\paragraph{Parametric dispersion model.}
For drop height $h$ (resolved via voxel-driven raycast, below):
\begin{equation}\label{eq:gravity_sigma}
\begin{aligned}
v_{\text{impact}} &= \sqrt{2 g h}, \qquad
  t_{\text{fall}}  = \sqrt{2 h / g}, \\
\sigma_{\text{bounce}} &= \varepsilon_{\text{rough}}\,v_{\text{impact}} \big/ \max(0.1,\,1-e), \\
\sigma_{\text{shape}}  &= r_{\text{obj}}\,\text{footprint\_factor}(\text{shape}), \\
\sigma_{\text{rel-v}}  &= \tfrac{1}{2}\,\|\bm v_{\text{rel,xy}}\|\,t_{\text{fall}}, \\
\sigma_{xy}^2 &= \sigma_{\text{bounce}}^2 + \sigma_{\text{shape}}^2 + \sigma_{\text{rel-v}}^2, \\
\sigma_{z} &= \tfrac{1}{2}\,r_{\text{obj}}\,(1 - e/2), \\
\sigma_{\text{yaw}} &= \pi\,\bigl(1 - \exp\!\bigl(-h(1-e^2)/r_{\text{obj}}\bigr)\bigr).
\end{aligned}
\end{equation}
Default $\varepsilon_{\text{rough}} = 5\!\cdot\!10^{-3}\,\text{m}\,/\,(\text{m s}^{-1})$;
footprint factor by shape: spherical $0.25$, cylindrical $0.50$, box $0.70$,
irregular $1.00$. Higher restitution increases $\sigma_{\text{bounce}}$ ---
physically, a rubber ball bounces several times each accruing scatter,
while a clay ball lands once and stays.

\paragraph{Voxel-driven $h$ branch.}
The downward raycast classifies the column into one of four states; each
adjusts $h$, landing $z$, and the $\sigma_z$ floor:
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item \texttt{hit\_occupied}: clean surface at $h^\star$;
  $h = h^\star$, $\sigma_z$ from \eqref{eq:gravity_sigma}.
\item \texttt{mixed\_unseen}: surface known but UNSEEN voxels above it;
  $h = h^\star$, $\sigma_z$ extra-inflated by $(h_u - h^\star)/2$ to reflect
  path uncertainty.
\item \texttt{all\_unseen}: surface unknown; $h = (h_u + h_{\text{floor}})/2$
  (\emph{midpoint}, not worst case), $\sigma_z = (h_u - h_{\text{floor}})/2$.
\item \texttt{all\_empty}: column observed empty all the way to the
  workspace floor; $h = h_{\text{floor}}$, $\sigma_z = r_{\text{obj}}$.
\end{itemize}
The midpoint choice for \texttt{all\_unseen} gives the EKF a useful prior
that subsequent observations can correct quickly, instead of pushing the
mean to an extreme that is unlikely to be right.

\paragraph{Trigger.}
The release predict fires once at $\{\textsc{Holding},\textsc{Releasing}\}\to
\textsc{Idle}$. A toggle flag (default on) gates the hook; with the voxel
grid unset, the hook short-circuits and the legacy static-$Q$-only release
behaviour is preserved exactly. Subsequent frames see the released track
as unobserved, so \eqref{eq:proc}'s keyframe freeze preserves the gravity
prior (in particular $\sigma_{yaw}=\pi$ does not pump $\sigma_{xy}$ via
cross-coupling).

\subsection*{II.12 Gripper phase FSM}

Four phases:
$\textsc{Idle}\to\textsc{Grasping}\to\textsc{Holding}\to\textsc{Releasing}\to\textsc{Idle}$.
\textbf{Project assumption:} the gripper begins fully open; ANY closing of
the fingers registers as a grasp action. The state machine therefore detects
``closed'' as a drop $\ge\!0.005$~m below the rolling open baseline (5~mm
captures any non-trivial finger motion while ignoring proprioception
jitter), in addition to the absolute thresholds. Drivers:
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Width threshold crossings with hysteresis (closed $0.025$~m, open
  $0.040$~m, drop-below-baseline $0.005$~m).
\item Stability test: last $5$ widths span $<\!0.01$~m.
\item Min-frames timer ($5$ frames) holds the transition phases through
  brief instability.
\end{itemize}
At the $\textsc{Idle}\to\textsc{Grasping}$ edge, the held oid is selected
by ranking detections by inside-gripper-volume containment fraction
($=$count\_inside/total\_pts); ties break by smallest mask area. A fallback
returns the nearest live track within $0.30$~m of $T_{we}$. Self-merges
propagate via an apply-merges hook: if the dropped id was the held seed,
it switches to the keep id.

\subsection*{II.13 Frame-by-frame algorithm}

\begin{algorithm}[H]
\caption{One frame of the Bernoulli-EKF tracker.}
\begin{algorithmic}[1]
\State \emph{Inputs}: $T_{wb}$, $T_{bc}$, $T_{bg}$, $w$, RGB, depth, detections.
\State $g \gets \textsc{GripperPhaseTracker}.\textsc{step}(w, \ldots)$
\State Maybe call relations backend (\S II.9 trigger).
\State $\mathcal{H} \gets \textsc{Expand}(g.\text{held},\,\mathcal{E}_t)\cap\text{live oids}$
\State $\mathcal{U} \gets \{o : \text{frames since obs}[o] > 0\}$ \Comment{keyframe-frozen set}
\State \textsc{predict\_static}(skip\,$=\!\mathcal{H}$,\;unobserved\,$=\!\mathcal{U}$);\,\textsc{rigid\_attachment\_predict}(each $o\in\mathcal{H}$)
\State \textsc{suppress\_subpart\_detections};\,\textsc{centroid\_cam\_from\_mask};\,gather detections
\State \textsc{hungarian\_associate} $\to$ matches, unmatched\_tracks, unmatched\_dets
\For{each match $(o, d)$}
  \If{$o = g.\text{held}$ and (held-prefilter or innov-clamp triggers)} skip update \EndIf
  \State \textsc{update\_observation\_centroid}; update $r_o$ via match log-likelihood
\EndFor
\For{each unmatched track $o$}
  \State Update $r_o$ via miss equation; floor at $r_{\text{held floor}}$ if $o = g.\text{held}$
\EndFor
\For{each unmatched detection $d$}
  \If{birth-admissible$(d)$ and not near-live$(d)$}
    \State Mint new oid, run ICP once, set $r_0$ per \eqref{eq:birth_r}
  \EndIf
\EndFor
\State Prune oids with $r_o < r_{\min}$ (skip held seed)
\State $\mathcal{P} \gets \{(\min(p,c),\max(p,c))\,:\,(p,c,\cdot)\in\mathcal{E}_t\}$
\State Self-merge pass with held$=g.\text{held}$ and protected$=\mathcal{P}$
\State $g$.apply-merges(merges); relation pipeline.remap-after-merges(merges)
\State Maybe gravity-predict (release transition: overwrite the just-released oid via \eqref{eq:gravity_sigma})
\end{algorithmic}
\end{algorithm}

\subsection*{II.14 Numerical defaults summary}

\begin{tabular}{@{}lll@{}}
$R_{cam}$ & $\mathrm{diag}((0.02\,\mathrm{m})^2)$ & centroid measurement \\
$Q_{\text{static}}$ & $1{\rm e}{-}6\to1{\rm e}{-}5\to1{\rm e}{-}8$ & schedule on frames since obs \\
$Q_{\text{manip seed}}$ (grasping/releasing) & $\mathrm{diag}((0.02)^2{\times}3,(0.10)^2{\times}3)$ & per frame \\
$Q_{\text{manip seed}}$ (holding) & $\mathrm{diag}(2.5{\rm e}{-}4{\times}3,1{\rm e}{-}3{\times}3)$ & per frame \\
$Q_{\text{manip transit}}$ & holding row above & transitive held members \\
$G_{\text{out},t}$ & $21.108$ & $\chi^2_3$ at $99.97\%$ \\
$G_{\text{in},t}$ & $7.815$ & $\chi^2_3$ at $95\%$ Huber inner \\
$\alpha_{\text{SAM2}}$ & $4.4$ & continuity bonus \\
$\max\_$residual & $0.30$~m & hard Euclidean residual cap (None disables) \\
$P_{\max}$ & none & covariance saturation cap (no cap default) \\
birth-min-dist & $0.05$~m & default near-live radius \\
birth-border-margin & $2$~px & image-edge margin \\
birth-pending-TTL & $30$ frames & pending-birth TTL ($\sim\!1$~s @ 30~Hz) \\
held-birth-radius & $0.25$~m & wider for the held oid \\
held-meas-radius & $0.25$~m & end-effector prefilter \\
held-meas-innov-max & $0.20$~m & innovation clamp \\
$r_{\min}$ & $10^{-3}$ & prune threshold \\
$r_{\text{held floor}}$ & $0.50$ & held existence floor \\
birth-confirm-$k$ & $3$ & tracker-side unmatched-streak count \\
birth-score-min & $0.20$ & rolling-max score floor \\
self-merge-trans & $0.05$~m & Euclidean merge gate \\
dedup-voxel-size & $0.02$~m & subpart suppression voxel size \\
dedup-containment-thresh & $0.8$ & subpart suppression threshold \\
$\beta$ (label penalty) & $0.0$ default & soft mode; production sets $6.0$ \\
$\gamma$ (score weight) & $0.0$ default & soft mode; production sets $2.0$ \\
closed/open width & $0.025\,\mathrm{m},\,0.040\,\mathrm{m}$ & FSM hysteresis \\
close-delta & $0.005$~m & drop-below-baseline = grasp \\
grasp-radius & $0.30$~m & nearest-track gate at grasp onset \\
EMA $\alpha$, threshold & $0.3,\,0.5$ & relation smoothing \\
relation-every-$n$-frames & $90$ & periodic backend tick \\
gravity-predict & on & one-shot release-time gravity hook \\
workspace-floor-$z$ & $-1.0$~m & column terminator for raycast-down \\
$\varepsilon_{\text{rough}}$ & $5\!\cdot\!10^{-3}$ & bounce-roughness coefficient \\
voxel grid (default) & $5$~cm at $5{\times}5{\times}3$~m & observability grid \\
$n_{\min}^{\text{hit}},\,n_{\min}^{\text{pass}}$ & $2,\,3$ & voxel hysteresis thresholds \\
\end{tabular}

\bigskip\hrule\bigskip

%=============================================================================
\section*{Part III --- Algorithm-to-Code Map}
\label{sec:map}

\sloppy

Each line below cites exactly one source-file path. Where one file hosts
multiple algorithmic items, the path repeats across multiple lines, each
line scoping a distinct role.

\subsection*{Public API and configuration}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Public five-method surface (\texttt{detect}, \texttt{step}, \texttt{get\_scene}, \texttt{get\_points}, \texttt{smooth}): \file{ekf\_tracker/api.py}.
\item Bernoulli/EKF/birth thresholds and gates (\S II.4, II.6 numerics): \file{ekf\_tracker/config.py}.
\item Live OWL+SAM2 streaming detection pipeline behind \texttt{detect}: \file{ekf\_tracker/perception\_pipeline.py}.
\end{itemize}

\subsection*{State}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Per-object Bernoulli-Gaussian belief and storage: \file{ekf\_tracker/state/gaussian\_state.py}.
\item Existence-probability arithmetic (\eqref{eq:bern_pred_r}, \eqref{eq:r_assoc}, \eqref{eq:r_miss}, \eqref{eq:birth_r}): \file{ekf\_tracker/state/bernoulli.py}.
\item SE(3) tangent algebra, Joseph form, Huber reweight, $Q$ schedule: \file{utils/ekf\_se3.py}.
\item Lift-to-base and information-sum primitives used by ICP and self-merge: \file{utils/object\_belief.py}.
\item Append-only per-track ICP-measurement chain (diagnostic, smoother substrate): \file{ekf\_tracker/state/obs\_chain.py}.
\item RBPF backend state (legacy fast tier; not the canonical path): \file{ekf\_tracker/state/rbpf\_state.py}.
\end{itemize}

\subsection*{Predict}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item World-static and rigid-attachment per-track predict (\eqref{eq:proc}, \eqref{eq:telescope}): \file{ekf\_tracker/state/gaussian\_state.py}.
\item Phase-aware $Q$ schedule and adjoint covariance transport: \file{utils/ekf\_se3.py}.
\item Step orchestration and held-set wiring: \file{ekf\_tracker/gaussian\_ekf\_tracker.py}.
\item Two-tier orchestrator gluing fast tier and slow tier: \file{ekf\_tracker/orchestrator\_gaussian.py}.
\end{itemize}

\subsection*{Detection conditioning}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Voxel-containment subpart suppression and label-histogram merge: \file{perception/det\_dedup.py}.
\item Mask cleaning, depth back-projection, centroid validity gate: \file{perception/icp\_pose.py}.
\end{itemize}

\subsection*{Association}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Hungarian solver, $\chi^2_3$ gate, residual cap, label penalty (\eqref{eq:cost_sam2}, \eqref{eq:hung}): \file{perception/association.py}.
\item Default values for $\alpha_{\text{SAM2}}$, $G_{\text{out}}$, $\beta$, $\gamma$ and the label-enforcement flag: \file{ekf\_tracker/config.py}.
\end{itemize}

\subsection*{Update}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Translation-only Joseph step on the centroid measurement (\S II.5): \file{ekf\_tracker/state/gaussian\_state.py}.
\item Full-$\SE$ Joseph form and IEKF inner iterations (used when ICP succeeds): \file{utils/object\_belief.py}.
\item EKF predict, EKF update on $\SE$, Huber reweight (\eqref{eq:huber_w}): \file{utils/ekf\_se3.py}.
\item World-frame composition $\mu_{wo},\Sigma_{wo}$ (\eqref{eq:world_compose}): \file{ekf\_tracker/state/gaussian\_state.py}.
\item World-frame pose-and-cov overwrite (release-hook primitive): \file{ekf\_tracker/state/gaussian\_state.py}.
\end{itemize}

\subsection*{Birth and prune}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Confirm-$k$ / score / border / TTL admission gate: \file{ekf\_tracker/birth\_gate.py}.
\item Near-live-track world-frame proximity gate: \file{perception/birth\_gating.py}.
\item Birth existence probability $r_0$ (\eqref{eq:birth_r}): \file{ekf\_tracker/state/bernoulli.py}.
\item One-shot ICP at admission (with mask cleaning and depth filter): \file{perception/icp\_pose.py}.
\end{itemize}

\subsection*{Self-merge}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Self-merge pass (candidate enumeration, sort, fuse) and merge-event emission: \file{ekf\_tracker/gaussian\_ekf\_tracker.py}.
\item Information-sum fusion primitive (\eqref{eq:self_merge_info}): \file{ekf\_tracker/state/gaussian\_state.py}.
\end{itemize}

\subsection*{Visibility}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Fibonacci-sphere sampling and depth-test visibility $p_v$ (\eqref{eq:pv}): \file{perception/visibility.py}.
\end{itemize}

\subsection*{Relations}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Trigger gate, backend dispatch, edge emission, drop-to-keep remap: \file{ekf\_tracker/relations/relation\_orchestrator.py}.
\item LLM and REST clients with disk-cached prompts/responses: \file{ekf\_tracker/relations/relation\_client.py}.
\item Per-edge EMA smoother (\S II.9): \file{ekf\_tracker/relations/relation\_filter.py}.
\item Held-set closure under in/on edges (\S II.10) and trigger predicate: \file{ekf\_tracker/relations/relation\_utils.py}.
\end{itemize}

\subsection*{Gripper FSM}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Four-phase FSM with hysteresis, stability, and min-frame timer: \file{utils/gripper\_state.py}.
\item Held-oid selection at the grasp edge (containment-fraction ranking): \file{ekf\_tracker/manipulation/grasp\_owner\_detector.py}.
\item Phase + held-set wiring into the tracker step: \file{ekf\_tracker/manipulation/gripper\_state\_inferrer.py}.
\item Gripper finger geometry primitives shared by FSM and visualizer: \file{utils/gripper\_geometry.py}.
\end{itemize}

\subsection*{Gravity-aware release}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Voxel observability grid and downward raycast classification: \file{perception/voxel\_observability.py}.
\item Per-label dynamics table $(e,\mu,\text{shape},r_{\text{obj}})$ with default fallback: \file{utils/object\_dynamics.py}.
\item Parametric dispersion model (\eqref{eq:gravity_sigma}) and landing-pose synthesis: \file{ekf\_tracker/manipulation/gravity\_predict.py}.
\item One-shot release-transition hook calling the overwrite primitive: \file{ekf\_tracker/api.py}.
\end{itemize}

\subsection*{Driver, harness, slow tier}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Canonical instrumented fast-tier driver with debug capture: \file{scripts/visualize\_ekf\_tracking.py}.
\item Fast-tier orchestration (per-frame step body, debug output): \file{ekf\_tracker/gaussian\_ekf\_tracker.py}.
\item Slow tier (factor-graph smoother) and its trigger schedule: \file{ekf\_tracker/factor\_graph.py}.
\item SLAM marginal interface and base-pose adapters: \file{utils/slam\_interface.py}.
\item Robot kinematics and end-effector geometry (Fetch baseline): \file{utils/robot\_models/fetch.py}.
\item Deep camera-pose refiner used only when explicitly enabled: \file{perception/camera\_pose\_refiner.py}.
\item Adaptive measurement-noise kernel for centroid likelihood: \file{perception/adaptive\_kernel.py}.
\end{itemize}

\subsection*{Tests and diagnostics}
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item Sole quantitative regression (parity against cached detections): \file{tests/parity\_apple\_drop.py}.
\item Live OWL+SAM2 detect $\to$ EKF step smoke test: \file{tests/integration/test\_detect\_step\_pipeline.py}.
\item Per-API mp4 renderer (5 separate videos, one per public method): \file{tests/diagnostics/render\_api\_videos.py}.
\item API demo with overlay panels (single-trajectory): \file{tests/demo\_apis.py}.
\end{itemize}

\fussy

\bigskip\hrule\bigskip

%=============================================================================
\section*{Part IV --- Reviewer Critique}
\label{sec:critique}

The audit underlying this part is summarised by category. Citations use
\texttt{file.py:line} for evidence. The critique is intended for a top-tier
robotics venue review and is intentionally blunt.

\subsection*{IV.A Single-trajectory hardcoding}

The runnable pipeline carries trajectory-specific anchors in its production
sources, not just in scripts:

\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item \texttt{ekf\_tracker/manipulation/gravity\_predict.py:46-47} sets the
  bounce-roughness coefficient $\varepsilon_{\text{rough}}=5\!\cdot\!10^{-3}$
  with the inline comment ``tune empirically on the apple\_drop trajectory''.
\item \texttt{ekf\_tracker/gaussian\_ekf\_tracker.py:82} justifies a held-set
  rule with a specific frame range of \texttt{apple\_drop fr 269-273}.
\item \texttt{ekf\_tracker/gaussian\_ekf\_tracker.py:924} cites a specific
  \texttt{apple\_in\_the\_tray} frame as the failure case the fix is keyed to.
\item \texttt{ekf\_tracker/state/gaussian\_state.py:612} bounds drift by
  citing ``apple\_in\_the\_tray oid 3, frames 290-330''.
\item \texttt{perception/icp\_pose.py:272} defends a heuristic with the
  apple\_in\_the\_tray frame-430 case.
\item \texttt{utils/fetch\_kinematics.py:17} calibrates camera-to-base
  rotation against \texttt{/tf} from the \texttt{apple\_in\_the\_tray} bag.
\item \texttt{utils/robot\_models/fetch.py:24, 53} pin empirical inside-pts
  validation to \texttt{apple\_in\_the\_tray frame 488}.
\end{itemize}

Comments-in-code that name a specific frame index of a specific bag are not
documentation; they are evidence that the algorithm was reverse-engineered
from one episode and was never re-validated on another. A new trajectory
that does not tickle these specific frame indices is not exercised by any
gate; behaviour on unobserved scene structure is undefined.

The vocabulary leaks reinforce this. The detection-pipeline tests and
demos all pass a fixed list shaped like
\texttt{["apple", "cup", "bottle", "cabinet", "bowl"]}
(\texttt{tests/demo\_apis.py:218}, \texttt{tests/diagnostics/render\_api\_videos.py:441-442},
\texttt{tests/integration/test\_detect\_step\_pipeline.py:106}). The
per-label dynamics table covers six labels (apple, milkbox, cola, cup, pot,
flowerpot) at \texttt{utils/object\_dynamics.py:69-76}; everything else is
silently mapped to a 5~cm spherical rigid body with $e{=}0.40,\,\mu{=}0.50$
through \texttt{utils/object\_dynamics.py:104-105}, with no warning emitted
when the fallback is taken.

Single-frame regression artefacts are also pervasive: \texttt{FRAME\_IDX
= 399} hardcoded for apple\_in\_the\_tray (\texttt{tests/diagnostics/dump\_llm\_relation\_io.py:34},
\texttt{tests/diagnostics/visualize\_llm\_relation\_frame.py:29}),
an entire test file pinned to that frame
(\texttt{tests/integration/test\_llm\_relation\_frame399.py}), and
\texttt{demo\_frame=281} hardcoded as the apple\_drop release boundary
(\texttt{tests/demo\_apis.py:208}). The sole quantitative regression in
the suite is \texttt{tests/parity\_apple\_drop.py}, a parity check against
cached detections on a single bag; it cannot fail unless cache and code
disagree, so it provides no signal about correctness on any other input.

\subsection*{IV.B Robot-platform and capture-rate hardcoding}

The pipeline silently assumes a Fetch parallel-jaw gripper running at
30~Hz. URDF finger-link offsets are hard-encoded into
\texttt{utils/robot\_models/fetch.py:11-14} from the public Fetch hardware
spec and then validated against a single bag-and-frame
(\texttt{utils/robot\_models/fetch.py:24, 53}). Width hysteresis values
in \texttt{utils/gripper\_state.py} (closed-width $25$~mm, open-width
$40$~mm, drop-below-baseline $5$~mm) are tuned for fruit-sized objects in
that gripper. Time constants -- birth-pending TTL of $30$ frames at
\texttt{ekf\_tracker/config.py:99}, FSM min-transition $5$ frames in
\texttt{utils/gripper\_state.py}, relation period $90$ frames in
\texttt{ekf\_tracker/relations/relation\_utils.py} -- are interpretable
only at $30$~Hz. None of these constants is exposed per-platform; deploying
on another robot or rate requires editing source.

\subsection*{IV.C Algorithmic narrowness}

\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item \textbf{Two predict regimes are too few.} World-static and
  rigid-attached miss sliding, rolling, in-flight, and any deformable
  regime. The gravity-aware branch is a one-shot replacement at the
  release transition (\texttt{ekf\_tracker/api.py:443}), not a process
  model. An object that rolls off a tray edge mid-trajectory falls back
  to static prediction with no further inflation.
\item \textbf{Translation-only Joseph + birth-time ICP} permanently
  forecloses post-birth rotation refinement
  (\texttt{ekf\_tracker/state/gaussian\_state.py:546-670}).
  Symmetric objects (cup, bowl) are unobservable in yaw forever; an object
  re-encountered at a novel pose accumulates uncertainty unboundedly with
  no mechanism for re-alignment.
\item \textbf{The keyframe freeze is correct only under stationary
  $T_{wb}$.} Loop-closure SLAM corrections retroactively shift past base
  poses; the frozen $\Sigma_{bo,k_f}$ is then no longer the marginal
  posterior. The diagnostic chain at
  \texttt{ekf\_tracker/state/obs\_chain.py} is the natural substrate for
  retroactive correction but is never read by the fast tier.
\item \textbf{Hard vs.\ soft label gate is two algorithms shipped under
  one name.} The default at \texttt{ekf\_tracker/config.py:59} is
  \texttt{enforce\_label\_match=True} (hard gate); the production driver
  at \texttt{scripts/visualize\_ekf\_tracking.py:2045-2068} overrides to
  soft mode with $\beta{=}6.0$, $\gamma{=}2.0$, and
  \texttt{enforce\_label\_match=False}. The paper-baseline gate breaks the
  moment the open-vocabulary detector flips ``apple''~$\leftrightarrow$~``tomato''
  on the same physical object across frames; the production gate is what
  is empirically validated. The two names should not be the same algorithm.
\item \textbf{Subpart suppression at $0.8$ voxel containment}
  (\texttt{perception/det\_dedup.py:127}) absorbs an item-shaped mask into
  a container-shaped mask whenever the item sits inside the container's
  silhouette --- exactly the configuration the system claims to track via
  in/on relations.
\item \textbf{Self-merge gate fixed at 5~cm Euclidean}
  (\texttt{ekf\_tracker/config.py:77}) has no scale-relative variant. A
  $1$~m cabinet collapses to a single track if SAM2 yields two masks; two
  adjacent items 5~cm apart in clutter survive arbitrarily.
\item \textbf{Held-set closure capped at 8 iterations}
  (\texttt{ekf\_tracker/relations/relation\_utils.py:30}). Containment
  chains longer than 8 (item~$\to$~bowl~$\to$~tray~$\to$~cart~$\to$~shelf~$\to\ldots$)
  are silently truncated.
\item \textbf{Visibility = Fibonacci sphere of $r_{\text{obj}}$}
  (\texttt{perception/visibility.py:59-87}) is wrong for flat (tray),
  elongated (broom), and non-convex objects. A flat tray seen edge-on
  registers near-zero $p_v$ even when fully visible from above; the
  geometric occlusion does not match the logical observability the model
  is supposed to capture.
\item \textbf{Object dynamics covers six labels} at
  \texttt{utils/object\_dynamics.py:69-76}; everything else is silently a
  $5$~cm spherical rigid body with $e{=}0.40,\,\mu{=}0.50$. Glass, foam,
  soft, articulated, and large objects all inherit this default with no
  warning.
\item \textbf{Workspace AABB fixed at $5\times 5\times 3$~m at 5~cm voxel
  size} (\texttt{perception/voxel\_observability.py}). Tabletop scenes
  waste 90\% of the grid; outdoor scenes truncate. Non-axis-aligned
  surfaces produce systematic biases in the raycast-down classification.
\item \textbf{Process-noise schedule and predict-static skip-cov is hidden
  cross-file coupling.} The non-monotonic $Q$ table at
  \texttt{utils/ekf\_se3.py:347-402} (idle $1{\rm e}{-}6\to$ unstable
  $1{\rm e}{-}5\to$ stable $1{\rm e}{-}8$) is only safe because
  \texttt{ekf\_tracker/state/gaussian\_state.py:267} skips covariance
  updates for the unobserved set. Either piece in isolation is wrong;
  refactoring either side breaks long-occlusion behaviour with no test
  signal.
\end{itemize}

\subsection*{IV.D Bayesian incoherences}

\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item \texttt{ekf\_tracker/config.py:104} sets $r_{\text{held floor}}=0.5$
  while \texttt{ekf\_tracker/config.py:52} sets $r_{\min}=10^{-3}$. A held
  track survives arbitrary missed observations regardless of evidence;
  this is not a posterior under the miss-likelihood model
  (\eqref{eq:r_miss}), it is a hold-out. The reported existence
  probability for the held seed therefore loses its frequentist meaning.
\item The translation-only update zeros the rotation cross-covariance
  rather than marginalising it
  (\texttt{ekf\_tracker/state/gaussian\_state.py:546-670}). Translation
  observations carry rotation information through the lever arm
  ($\Ad(T)$ has off-diagonal $[t]_\times R$ blocks); that information is
  discarded by construction. The honest rule is to set $R_{rot}=\infty$
  and run the full Joseph form on the joint covariance, which would
  yield the same translation block but a properly marginalised rotation
  block. The current shortcut is a documentation hazard.
\item Subpart suppression merges label histograms but discards the
  absorbed track's existence posterior
  (\texttt{perception/det\_dedup.py:107-170}). A high-$r$ small object
  contained inside a low-$r$ container loses its existence weight on
  absorption.
\item The EKF likelihood used for $r_{\text{assoc}}$ at
  \eqref{eq:r_assoc} treats the $\chi^2$ rejection as a maximum-likelihood
  decision and ignores the gating prior on assignment. The Hungarian
  decision is therefore not coherent with the existence update under any
  joint posterior on $(\sigma,r)$.
\end{itemize}

\subsection*{IV.E Evaluation vacuum}

The only quantitative test in the suite is
\texttt{tests/parity\_apple\_drop.py} --- a parity match against cached
detections on a single bag. There is no held-out scene, no pose-error
ground-truth metric, no frame-rate stress test, no clutter scaling, no
multi-robot test, no comparison against any baseline tracker. The phrase
``verified to machine precision on apple\_drop oid=1, fso=224'' that
appears in earlier documentation is a numerical sanity check on the
algebraic identity \eqref{eq:telescope}, not a validation of the algorithm
against external truth.

This documentation itself shows code/doc drift: prior to this rewrite, the
LaTeX source referenced a \texttt{pose\_update/} directory that has been
refactored out of the repository entirely. A system whose written
description does not match its source tree is, by construction, not
audited end-to-end.

\subsection*{IV.F Top-3 worst single-scenario assumptions}

\begin{enumerate}[leftmargin=1.5em,itemsep=2pt,topsep=2pt]
\item $r_{\text{held floor}}=0.5$ (\texttt{ekf\_tracker/config.py:104}).
  Held tracks are pinned alive regardless of evidence. This silently keeps
  phantom held-object identities alive across long occlusions and turns
  the existence probability into a meaningless number for the held seed.
\item \texttt{LABEL\_DYNAMICS\_TABLE} covers six fruit-and-container
  labels (\texttt{utils/object\_dynamics.py:69-76}). Every other label is
  silently mapped to a $5$~cm spherical rigid body with $e{=}0.40,\,
  \mu{=}0.50$. The gravity-aware release predict therefore degrades to
  a fixed-shape, fixed-restitution prior outside the validated label set
  with no caller signal.
\item Hidden coupling between the $Q$-schedule
  (\texttt{utils/ekf\_se3.py:347-402}) and the predict-static unobserved
  branch (\texttt{ekf\_tracker/state/gaussian\_state.py:267}). The
  non-monotonic $Q$ table is only safe because the unobserved branch
  skips the covariance update; refactoring either side in isolation
  breaks long-occlusion tracking with no test signal --- and the only
  test in the suite is parity on a cached bag, so the regression cannot
  surface.
\end{enumerate}

\bigskip\hrule\bigskip

%=============================================================================
\section*{Part V --- Per-Component Justification Audit}
\label{sec:audit}

For each algorithmic component, exactly one of the following is given:
(i) a published \textbf{Reference} that directly supports the design
choice, (ii) an undeniable \textbf{Principled} reason
(algebraic, information-theoretic, or Bayesian), or
(iii) a \textcolor{red}{\textbf{REDUNDANT}} flag indicating the
component is unjustified relative to its complexity, with a recommended
replacement.

\subsection*{V.1 State and storage}

\paragraph{Per-object Bernoulli--Gaussian belief.}
\REFERENCE{} Vo \& Vo, ``The Cardinality Balanced Multi-Target
Multi-Bernoulli Filter and Its Implementations'' (IEEE TSP, 2009);
Mahler, \emph{Advances in Statistical Multisource-Multitarget
Information Fusion} (Artech House, 2014), Ch.~7--8.

\paragraph{Body-frame storage with SLAM marginal decoupled.}
\PRINCIPLED{} Storing $\mu_{bo},\Sigma_{bo}$ in the base frame leaves
the recursion independent of $T_{wb}$; the world view
$T_{wo}=T_{wb}\,\mu_{bo}$ is the deterministic composition. SLAM
updates no longer touch the per-object recursion; per-object updates
no longer touch the SLAM marginal. Standard lever-arm decoupling, see
Barfoot, \emph{State Estimation for Robotics} (2017), \S 6--\S 8.

\paragraph{Right-trivialised SE(3) tangent (translation-then-rotation).}
\REFERENCE{} Sol\`a, Deray, Atchuthan, ``A micro Lie theory for state
estimation in robotics'' (arXiv:1812.01537, 2018). The choice of
right-trivialisation and ordering convention is standard for SE(3)
Kalman filtering.

\paragraph{Dual-backend abstraction (Gaussian + RBPF).}
\REDUNDANT{} The repository ships both backends, but the RBPF tier is
acknowledged buggy and the canonical instrumented driver
(\texttt{scripts/visualize\_ekf\_tracking.py:329}) exercises only the
Gaussian backend. The shared interface in
\texttt{utils/base\_pose\_backend.py} therefore duplicates every
primitive (\texttt{update\_observation},
\texttt{update\_observation\_centroid}, \texttt{predict\_static},
\texttt{predict\_static\_all}) into \texttt{rbpf\_state.py} that no
production path consumes.
\textbf{Fix:} delete \texttt{ekf\_tracker/state/rbpf\_state.py} and
\texttt{utils/base\_pose\_backend.py}; collapse the API surface to the
Gaussian backend.

\paragraph{\texttt{predict\_static\_all} alongside \texttt{predict\_static}.}
\REDUNDANT{} \texttt{predict\_static\_all} is a one-line wrapper that
forwards its arguments unchanged to \texttt{predict\_static}
(\texttt{ekf\_tracker/state/gaussian\_state.py:703-720}). It exists
only to satisfy the dual-backend interface above.
\textbf{Fix:} delete with the dual-backend abstraction.

\subsection*{V.2 Predict}

\paragraph{World-static motion ($\mu_{bo,k}=u_k\mu_{bo,k-1}$).}
\PRINCIPLED{} Direct algebraic consequence of
$T_{wb,k}\mu_{bo,k}=T_{wb,k-1}\mu_{bo,k-1}$. Identity arithmetic; no
external reference needed.

\paragraph{Rigid-attached motion ($\mu_{bo,k}=\delta_{bg,k}\mu_{bo,k-1}$).}
\PRINCIPLED{} Direct algebraic consequence of the rigid-attachment
assumption (object pose constant in gripper frame).

\paragraph{Adjoint covariance transport.}
\REFERENCE{} Bourmaud, M\'egret, Arnaudon, Giremus, ``Continuous-Discrete
Extended Kalman Filter on Matrix Lie Groups Using Concentrated
Gaussian Distributions'' (J.~Math.~Imag.~Vis., 2014); Forster, Carlone,
Dellaert, Scaramuzza, ``On-Manifold Preintegration for Real-Time
Visual-Inertial Odometry'' (IEEE T-RO, 2017).

\paragraph{Keyframe-freeze branch (skip cov update for unobserved tracks).}
\REDUNDANT{} The cross-coupling pumping that this freeze patches
($[t(u_k)]_\times P_{rr}[t(u_k)]_\times^\top$ in the Ad-conjugate) is
itself an artefact of an over-large process noise on rotation. The
freeze is non-Bayesian: under it, the recursion's posterior at frame
$k$ is no longer the marginal of any joint posterior, and any
retroactive SLAM correction (loop closure) is silently inconsistent
with the frozen $\Sigma_{bo,k_f}$.
\textbf{Fix:} set $Q_{\text{static}}=0$ for genuinely static objects
(no random walk, no scheduled inflation) and run the recursion
uniformly. The cross-coupling is then driven only by genuine motion.

\paragraph{$Q$ schedule on \texttt{frames\_since\_obs} ($1{\rm e}{-}6\to1{\rm e}{-}5\to1{\rm e}{-}8$).}
\REDUNDANT{} Adaptive $Q$ on observation age has no Bayesian
justification. The non-monotonic ramp (small idle / larger ``unstable''
/ smaller ``stable'') is tuned to mimic the keyframe-freeze behaviour
approximately; it is a workaround for a workaround. Maybeck,
\emph{Stochastic Models, Estimation, and Control} (1979), \S 1.5: $Q$
should reflect physical motion uncertainty, not estimator state.
\textbf{Fix:} one $Q$ per regime (static / manipulation), no schedule.

\paragraph{Covariance saturation $\Phi$ (trace cap on $P$).}
\REDUNDANT{} The default \texttt{P\_max=None}
(\texttt{ekf\_tracker/config.py:55}) means the cap is dead code in
production. Trace-clipping violates posterior optimality and
consistency (Maybeck \S 1.5).
\textbf{Fix:} delete \texttt{saturate\_covariance} and the
\texttt{P\_max} kwarg. If covariance grows unboundedly, repair the
predict / observation models, do not clip.

\subsection*{V.3 Detection conditioning}

\paragraph{Voxel-containment subpart suppression.}
\PRINCIPLED{} 3D analogue of 2D non-maximum suppression. A small
mask whose voxel set is mostly inside a larger mask's voxel set is,
under rigid-body assumptions, almost certainly an over-segmentation of
the larger mask. Compare 2D NMS in Neubeck \& Van Gool, ``Efficient
Non-Maximum Suppression'' (ICPR, 2006).

\paragraph{Centroid validity gate.}
\PRINCIPLED{} A mask without valid-depth pixels carries zero 3D
information; dropping it is the only honest behaviour.

\subsection*{V.4 Association}

\paragraph{Hungarian assignment.}
\REFERENCE{} Munkres, ``Algorithms for the Assignment and
Transportation Problems'' (J.~SIAM, 1957); Bar-Shalom \& Fortmann,
\emph{Tracking and Data Association} (Academic Press, 1988), \S 5.

\paragraph{$\chi^2_3$ probabilistic gate at $G_{\text{out},t}=21.108$.}
\REFERENCE{} Bar-Shalom, Li, Kirubarajan, \emph{Estimation with
Applications to Tracking and Navigation} (Wiley, 2001), \S 3.3.
$\chi^2_n$ tail at $99.97\%$ is a standard MIPDA outer-gate.

\paragraph{Translation-only Mahalanobis (assume $R_{rot}=\infty$).}
\PRINCIPLED{} Equivalent to marginalising rotation out of the joint
$\chi^2_6$. The translation-block residual under linear-Gaussian
assumptions is exactly $\chi^2_3$.

\paragraph{Hard Euclidean residual cap (\texttt{max\_residual\_m}=$0.30$).}
\REDUNDANT{} This second gate is admitted in the source comment to
catch the ``inflated-covariance pathology'' --- it patches a failure
of the $\chi^2$ gate caused by the keyframe-freeze and $Q$-schedule
machinery (V.2). Two gates with one principle is one too many.
\textbf{Fix:} repair $\Sigma_{bo}$ first (V.2); then remove the cap and
let the $\chi^2$ gate be the sole probabilistic admission criterion.

\paragraph{SAM2 continuity bonus $\alpha=4.4$.}
\REDUNDANT{} The mechanism (cost discount for matched perception ids)
is plausible but the value $4.4$ has no principled derivation. The
post-hoc ``$\chi^2_3$ at $78\%$'' interpretation is unmotivated.
\textbf{Fix:} replace with a likelihood factor
$\log p(\tau_o=\mathrm{pid}_d)$ derived from SAM2's empirical
id-stability statistics, or remove and rely on the $\chi^2$ gate
alone. Encoding identity continuity as a likelihood places it in the
same probabilistic framework as the $\chi^2$ gate.

\paragraph{Hard label gate (default \texttt{enforce\_label\_match=True}).}
\REDUNDANT{} The hard gate is the $\beta=\infty$ limit of the soft
penalty (next item). Two cost-function formulations for one purpose;
the production driver overrides the default to soft mode, so the hard
mode is a non-default that no shipped code path actually exercises in
regression.
\textbf{Fix:} keep the soft formulation only. Set $\beta$ large
(but finite) when behaviour-equivalent to ``hard'' is wanted.

\paragraph{Soft label penalty $\beta$ and score weight $\gamma$.}
\PRINCIPLED{} Augmenting the assignment cost with a label-mismatch
penalty and a low-score penalty is a likelihood-style construction.
The specific values ($\beta=6.0,\,\gamma=2.0$ in the production driver)
are unjustified constants but the mechanism is sound, provided one
formulation (this one) is canonical.

\subsection*{V.5 Update}

\paragraph{Joseph form on full SE(3).}
\REFERENCE{} Bar-Shalom, Li, Kirubarajan (2001), \S 5.3; Maybeck
(1979), \S 6.2. PSD by construction.

\paragraph{Translation-only Joseph with $P_{rt}^{\text{post}}=0$.}
\REDUNDANT{} The shortcut zeros the rotation cross-covariance instead
of marginalising it. Setting $R_{rot}=\infty$ in the full Joseph
update yields the same translation block \emph{and} a properly
marginalised rotation block via the Schur complement (Maybeck,
\S 6.2; Crassidis \& Junkins, \emph{Optimal Estimation of Dynamic
Systems}, 2011, \S 3.5). The current path discards the
lever-arm-coupled rotation information that translation observations
actually carry through the off-diagonal $[t]_\times R$ blocks of
$\Ad(T)$.
\textbf{Fix:} use the full Joseph form with $R_{rot}$ a large finite
number. The rotation block then updates correctly and the code path
is unified with the ICP update.

\paragraph{Held-track prefilters (\texttt{held\_meas\_radius\_m}=$0.25$,
\texttt{held\_meas\_innov\_max\_m}=$0.20$).}
\REDUNDANT{} Two ad-hoc Euclidean gates added on top of the $\chi^2$
gate, the residual cap, and the held-mode $Q$. Four overlapping
mechanisms guard one update. The principled mechanism is the
held-mode $Q$ schedule (large during grasping/releasing, small during
holding) which already encodes the physically expected innovation.
\textbf{Fix:} remove the prefilters. Trust the $Q$ schedule and the
$\chi^2$ gate.

\paragraph{One-shot ICP at admission.}
\PRINCIPLED{} Pragmatic compute budget: ICP per frame is expensive,
ICP once at birth gives a reasonable rotation prior. The known
weakness (no rotation refinement post-admission) is documented and
acknowledged. Reference: Besl \& McKay, ``A Method for Registration
of 3-D Shapes'' (IEEE PAMI, 1992).

\paragraph{Three ICP backends (\texttt{centroid} / \texttt{icp\_chain}
/ \texttt{icp\_anchor}).}
\REDUNDANT{} Default is \texttt{icp\_chain}; the other two are
documented as ``what the original integration test used'' and ``strict
decomposition'' respectively, and neither is exercised by any
production test.
\textbf{Fix:} keep \texttt{icp\_chain}; move \texttt{centroid} and
\texttt{icp\_anchor} to a dedicated \texttt{tests/icp\_backends/}
fixture, or delete.

\paragraph{Huber inner / outer reweight.}
\REFERENCE{} Huber, ``Robust Estimation of a Location Parameter''
(Annals of Math.~Stat., 1964); Zhang, ``Parameter Estimation
Techniques: A Tutorial'' (Image \& Vision Computing, 1997).

\paragraph{Adaptive Barron robust loss (slow tier only).}
\REFERENCE{} Chebrolu, Magistri, L\"abe, Stachniss, ``Adaptive Robust
Kernels for Non-Linear Least Squares Problems'' (IEEE RA-L, 2021).
Cited explicitly in the module header
\texttt{perception/adaptive\_kernel.py:2}.

\paragraph{World-frame composition $\mu_{wo},\Sigma_{wo}$
(\eqref{eq:world_compose}).}
\PRINCIPLED{} First-order propagation of the joint
$(T_{wb},\mu_{bo})$ uncertainty under deterministic composition,
treating $T_{wb}$ and $\mu_{bo}$ as independent. Standard lever-arm
arithmetic; see Barfoot (2017), \S 7.5.

\paragraph{Dual update primitives (\texttt{update\_observation} vs
\texttt{update\_observation\_centroid}).}
\REDUNDANT{} Two implementations of the same primitive: full-SE(3)
Joseph and translation-only Joseph. The centroid form is the special
case of the full form with $R_{rot}\to\infty$. The duplication is
also mirrored into \texttt{rbpf\_state.py}, giving four implementations
of one primitive.
\textbf{Fix:} implement once in the full form; the centroid case is
recovered by passing $R_{rot}$ large.

\subsection*{V.6 Birth and prune}

\paragraph{Bernoulli existence updates ($r$-predict, $r$-assoc, $r$-miss).}
\REFERENCE{} Mahler (2014), Ch.~7--8; Vo, Vo, Hoseinnezhad, ``Robust
Multi-Bernoulli Filtering'' (IEEE J.~STSP, 2013).

\paragraph{Birth existence $r_0=\lambda_b\hat s/(\lambda_b\hat s+\lambda_c)$.}
\REFERENCE{} Vo \& Vo (IEEE TSP, 2009). Standard CB-MeMBer birth.

\paragraph{Confirm-$k$ M-of-N gate.}
\REFERENCE{} Bar-Shalom \& Fortmann (1988), \S 3. M-out-of-N
detection is classical track-confirmation.

\paragraph{Confirm-$k$=$3$, score-floor=$0.20$, border-margin=$2$\,px,
TTL=$30$\,frames, near-live=$0.05$\,m.}
\REDUNDANT{} Magic numbers tuned for one trajectory's image
resolution, capture rate, and object scale. Border $2$\,px assumes a
$640\times 480$ stream; TTL=$30$ frames assumes $30$\,Hz capture;
near-live=$5$\,cm assumes fruit-sized objects.
\textbf{Fix:} replace with relative gates --- border $=1\%$ of image
width, TTL $=N$ seconds at any rate, near-live $=k\sigma_t$ on the
joint translation cov.

\paragraph{$r_{\min}=10^{-3}$ prune threshold.}
\PRINCIPLED{} Standard Bernoulli filter pruning on a memory /
precision tradeoff. The order of magnitude follows directly from the
floating-point sensitivity of \eqref{eq:r_assoc}.

\paragraph{$r_{\text{held floor}}=0.5$.}
\REDUNDANT{} Mathematically incoherent. The held seed's posterior
$r$ should be governed by \eqref{eq:r_miss}; if the model says $r$
should fall below $0.5$ given the miss-likelihood evidence, it should
fall. Pinning $r$ above $0.5$ converts it into a non-probability.
\textbf{Fix:} remove. If a held seed must survive long occlusions,
encode the prior via $p_d \to 0$ during full occlusion (which is
exactly what the visibility module is for) rather than by floor-pinning
the posterior.

\subsection*{V.7 Self-merge}

\paragraph{Information-sum fusion of two Gaussians.}
\REFERENCE{} Mutambara, \emph{Decentralized Estimation and Control
for Multisensor Systems} (CRC Press, 1998), \S 4.3. Approximate
(independence assumed) but the standard track-to-track fusion form.

\paragraph{Same-label gate at fixed $5$\,cm Euclidean
(\texttt{self\_merge\_trans\_m}=$0.05$).}
\REDUNDANT{} Fixed-distance, no scale-relative variant. Two adjacent
fruit at $5$\,cm survive arbitrarily; two SAM-split masks of one
cabinet collapse arbitrarily.
\textbf{Fix:} replace with $d^2_t \le \chi^2_3(0.95)=7.815$ on the
joint translation cov $P_{tt,o}+P_{tt,d}$.

\paragraph{Protected-pairs gate (relations-derived $\mathcal P$).}
\PRINCIPLED{} If the scene graph asserts two oids are physically
distinct (e.g.\ apple on tray), they must not collapse. Direct logical
consequence of the relation semantics.

\subsection*{V.8 Visibility}

\paragraph{Continuous $p_v$ in miss-frame likelihood
($\tilde p_d=p_d\cdot p_v$).}
\PRINCIPLED{} Standard Bernoulli filter convention under known
visibility (Mahler, 2014, \S 7.6).

\paragraph{Fibonacci-sphere visibility predicate.}
\REDUNDANT{} The principled visibility test is a depth-test against
the per-object TSDF / SDF (Newcombe et al., ``KinectFusion: Real-Time
Dense Surface Mapping and Tracking,'' ISMAR 2011). The spherical
fallback is wrong for any non-spherical object: a flat tray seen
edge-on registers near-zero $p_v$ even when fully visible from above.
\textbf{Fix:} implement TSDF-projection visibility per-object, or
demote the visibility module to a binary in-frustum test only.

\subsection*{V.9 Relations}

\paragraph{Per-edge EMA smoothing ($\alpha=0.3$, threshold $0.5$).}
\PRINCIPLED{} First-order low-pass to reject single-frame backend
flicker. The mechanism (any digital-filter text) is sound; the
specific $\alpha$ is a time-constant choice but does not affect
correctness.

\paragraph{Trigger gate (event triggers $+$ every-$90$-frame heartbeat).}
\REDUNDANT{} The event-trigger set (first call, release transitions,
new-oid arrivals) covers every state change that could alter the scene
graph. The $90$-frame periodic backstop is a hedge against unspecified
event-coverage gaps.
\textbf{Fix:} drop the heartbeat. If the event set under-covers, audit
the event set instead of patching with periodic re-fires.

\paragraph{Convention swap (parent=supported, child=supporter, against
the LLM's $j$-on-$i$).}
\REDUNDANT{} Exists to fix a sign-error regression. Two conventions
mixed in one codebase, with a swap at every emission point.
\textbf{Fix:} canonicalise once at the LLM client boundary --- the
output convention should match held-set expansion's parent/child
semantics directly. No swap.

\subsection*{V.10 Held-set expansion}

\paragraph{Closure under in/on edges.}
\PRINCIPLED{} Held-set semantics: anything supported by the held seed
is also held. Direct logical consequence.

\paragraph{Iteration cap of $8$.}
\REDUNDANT{} A fixed cap on a fixpoint iteration cannot accommodate
arbitrary chain depth. Containment chains in real scenes can exceed
$8$ (item-on-bowl-on-tray-on-cart-on-shelf-\ldots).
\textbf{Fix:} run the closure to convergence with a
\texttt{while changed:} loop. Worst case is $O(N\cdot|\mathcal E|)$,
still trivial.

\subsection*{V.11 Gravity-aware release}

\paragraph{Voxel observability grid (3-state with hit / pass hysteresis).}
\REFERENCE{} Hornung, Wurm, Bennewitz, Stachniss, Burgard, ``OctoMap:
An Efficient Probabilistic 3D Mapping Framework Based on Octrees''
(Auton.~Robots, 2013). The 3-state model is a coarse simplification
of OctoMap's log-odds occupancy update with explicit hysteresis instead
of log-odds clamping.

\paragraph{Voxel-driven $h$ branches \texttt{hit\_occupied} and
\texttt{all\_empty}.}
\PRINCIPLED{} Both have unambiguous geometric semantics: a clean
surface terminates the column at a known $h^\star$; a fully-empty
column terminates at the workspace floor.

\paragraph{Voxel-driven $h$ branches \texttt{mixed\_unseen} extra
inflation and \texttt{all\_unseen} midpoint heuristic.}
\REDUNDANT{} The midpoint choice for \texttt{all\_unseen} is an
arbitrary compromise between worst-case and best-case; the
\texttt{mixed\_unseen} ``path-uncertainty'' inflation $(h_u-h^\star)/2$
has no derivation.
\textbf{Fix:} for \texttt{all\_unseen}, return a flat uninformative
prior over the reachable depth range (uniform distribution lifted to
the EKF as a Gaussian with matching first two moments). Drop the
\texttt{mixed\_unseen} extra inflation.

\paragraph{Object dynamics table $(e,\mu,\text{shape},r_{\text{obj}})$
for six labels (\texttt{apple, milkbox, cola, cup, pot, flowerpot}).}
\REDUNDANT{} Six labels, values cited as ``from agricultural /
engineering literature'' with no per-label paper, plus a silent fallback
for every other label. Brittle to ontology drift; the
dynamics-conditioned $\sigma$ field of the gravity prior contributes at
most one frame of covariance and is then either confirmed or overwritten
by subsequent observations.
\textbf{Fix:} replace the table with one conservative prior
$P_{\text{drop}}=\mathrm{diag}((10\,\text{cm})^2{\times}3,\,(\pi/2)^2{\times}3)$
(or similar). Let the EKF re-converge from observations.

\paragraph{Parametric dispersion model
(\eqref{eq:gravity_sigma}: $\sigma_{xy},\sigma_z,\sigma_{\text{yaw}}$
in terms of $h$, $e$, $r_{\text{obj}}$, $\bm v_{\text{rel}}$).}
\REDUNDANT{} The functional forms are invented: the bounce term
$\sigma_{\text{bounce}}\propto v_{\text{impact}}/(1-e)$ has no
impact-mechanics citation; the yaw term
$\sigma_{\text{yaw}}=\pi(1-\exp(-h(1-e^2)/r_{\text{obj}}))$ is fitted,
not derived. The constant $\varepsilon_{\text{rough}}=5\!\cdot\!10^{-3}$
carries the source comment ``tune empirically on the apple\_drop
trajectory'' --- an explicit admission that the value is fitted on one
trajectory.
\textbf{Fix:} delete the dispersion model. Replace with the constant
prior above. The EKF posterior after two or three subsequent frames
of valid detection is indistinguishable from the dispersion-model
output, and this version has no scenario-tied tuning.

\paragraph{One-shot release-trigger
(\textsc{Releasing}$\to$\textsc{Idle}, overwrite EKF state once).}
\REDUNDANT{} A proper Bayesian filter has a
\textsc{Falling}/\textsc{Settling} regime that runs every frame
between holding and re-observation. The one-shot is a hack to avoid
implementing a continuous ballistic predict.
\textbf{Fix:} add a \textsc{Falling} predict to the predict step
(ballistic motion under gravity until the predicted impact frame, then
a transition to the constant-prior state); remove the one-shot
overwrite.

\paragraph{Workspace AABB fixed at $5\times 5\times 3$\,m, $5$\,cm voxel.}
\REDUNDANT{} Fixed-size hardcoded grid wastes $90\%$ of the volume on
tabletop scenes and truncates outdoor scenes.
\textbf{Fix:} derive the AABB from the SLAM map's bounding box plus a
margin; pick voxel size from $r_{\text{obj}}$ (e.g.~$r_{\text{obj}}/4$)
at allocation time.

\subsection*{V.12 Gripper FSM}

\paragraph{Four-phase FSM with width hysteresis and stability test.}
\PRINCIPLED{} Standard FSM with debouncing
(\AA str\"om \& Murray, \emph{Feedback Systems}, Princeton 2008,
\S 9). Hysteresis is the textbook anti-flicker mechanism for a
threshold-driven discrete state.

\paragraph{Width thresholds (closed=$25$\,mm, open=$40$\,mm,
delta=$5$\,mm, min-frames=$5$).}
\REDUNDANT{} Tied to one robot's gripper geometry and one frame rate.
\textbf{Fix:} derive thresholds from the URDF maximum width as
percentages (closed $=25\%$ of max, open $=50\%$); convert
\texttt{min\_transition\_frames} to a time constant ($150$\,ms) and
divide by the frame rate.

\paragraph{Held-oid selection by inside-gripper containment fraction.}
\PRINCIPLED{} The mask whose 3D points are most-contained in the
gripper finger volume is, by direct geometric inference, the most
likely held object. The smallest-mask-area tiebreak is also
principled (small held items are harder to over-segment than large
background items).

\subsection*{V.13 Codebase-level redundancies}

\paragraph{\texttt{perception/camera\_pose\_refiner.py} (337 LoC).}
\REDUNDANT{} Imported only by \texttt{scripts/data\_demo.py} and
\texttt{scripts/realtime\_app.py}, both legacy. Imports
\texttt{heuristic\_tracker.scene\_object} from the heuristic tracker
that the EKF replaced.
\textbf{Fix:} delete the module and its two callers.

\paragraph{Slow tier in \texttt{ekf\_tracker/factor\_graph.py} (when
not promoted).}
\PRINCIPLED{} The slow tier is a factor-graph smoother that fires only
when explicitly promoted. As an out-of-band module this is
acceptable; it is not on the production fast-tier critical path.

\paragraph{\texttt{ekf\_tracker/manipulation/gripper\_state\_inferrer.py}
($45$-line shim).}
\PRINCIPLED{} Driver-side wrapper that adapts
\texttt{utils/gripper\_state.py:GripperPhaseTracker} to the
tracker-coupling shape used by the EKF. Acceptable separation of
concerns; the FSM logic itself is not duplicated.

\subsection*{V.14 Tally and recommendation}

Of the $58$ components catalogued above:
\begin{itemize}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item $\textbf{13}$ are \textbf{Reference}-justified (Bernoulli MTT
  literature, Bar-Shalom and Mahler, Joseph form, Munkres, Huber,
  Barron / Chebrolu, OctoMap).
\item $\textbf{19}$ are \textbf{Principled} (algebraic, geometric, or
  Bayesian consequences with no useful citation beyond the obvious).
\item $\textcolor{red}{\textbf{26}}$ are
  \textcolor{red}{\textbf{REDUNDANT}}.
\end{itemize}

The redundancies cluster around four loci:
\begin{enumerate}[leftmargin=1.5em,itemsep=0pt,topsep=2pt]
\item \textbf{Predict patches (V.2):} the keyframe-freeze /
  $Q$-schedule / saturation triple together patch a problem that would
  not exist under an honest fixed $Q$.
\item \textbf{Update overlap (V.4--V.5):} four overlapping admission
  gates (\texttt{max\_residual\_m}, \texttt{held\_meas\_radius\_m},
  \texttt{held\_meas\_innov\_max\_m}, plus the $\chi^2$ gate) and a
  trans-only Joseph that discards rotation information instead of
  marginalising it.
\item \textbf{Gravity-predict subsystem (V.11):} an entire 200+ LoC
  subsystem ($\sigma_{xy},\sigma_z,\sigma_{\text{yaw}}$, label-conditioned
  restitution table, four-state column classification) that contributes
  one frame of covariance and is empirically tuned on
  \texttt{apple\_drop}.
\item \textbf{Magic-number ribbon (V.6 / V.12):} fixed birth, FSM,
  workspace, and self-merge constants that tie the system to one
  trajectory's image resolution, capture rate, robot platform, and
  object scale.
\end{enumerate}
Removing the $26$ redundant items is estimated to simplify the codebase
by $30$--$40\%$ (the gravity-predict subsystem alone is a quarter of
the redundancy mass) without changing any production-validated
behaviour, since none of the redundant items is exercised by an
external regression test that is not equally satisfied by its proposed
replacement.

\end{document}
